---
title: "Behavior of Twitter Bot in California Fires"
author: "Takeshi Oda"
date: "8/12/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggformula)
library(tm)
library(wordcloud)

```

## Parameter Settings

```{r parameter}
## Constant Variables ##
threshhold = 0.5
```

## Data Loading
```{r load, echo=TRUE}
tweets <- fread("data/CaliforniaFires_Tweet_Stats.csv", 
                encoding="UTF-8",
                colClasses = c("numeric",
                               "character",
                               "character",
                               "character",
                               "numeric",   #bot_probability
                               "integer",   #row_num
                               "integer",   #num_words
                               "integer",   #num_question
                               "integer",   #num_exclamation
                               "integer",   #num_digit_screen_name
                               "integer",   #num_political_word
                               "integer",   #num_environmental_word
                               "factor",    #include_retweet
                               "integer"    #num_hashtag
                               ))
```

## Exploratory Data Analysis
Bot probability was plotted on boxplot and percentiles are identified.
According to a prior research, it is estimated between 9% and 15% of active twitter users are bot(Varol, Ferrara, Davis, Menczer & Flammini, 2017).
If I apply this percentage into our dataset, the lower boundary of bot probability will be 0.125145748(85% percentile).
In this analysis,I chose more conservative threshold 0.5 to decide Bot account.  
```{r plotting, echo=TRUE}

summary(tweets)
gf_boxplot(~bot_probability, data=tweets, title="Bot Probability of twitter posting")
quantile(tweets$bot_probability, c(0,0.1,0.25,0.5,0.75,0.85,0.9,1))
tweets <- tweets %>% mutate(is_bot = ifelse(bot_probability > threshhold, "Bot", "Non-Bot"))

#Retweet
gf_histogram(~retweet, data=tweets, title="Number of retweet") 
ggplot(data=tweets, aes(x=log(retweet))) + facet_grid(. ~is_bot) + geom_histogram() + labs(title="Number of retweet(log transformation)")
#Number of question mark
ggplot(data=tweets, aes(x=num_question)) + facet_grid(. ~is_bot) + geom_bar() + geom_text(stat='count', aes(label=..count..), vjust=-1, size=3) + labs(title="Number of question mark")
#Number of exclamation mark
ggplot(data=tweets, aes(x=num_exclamation)) + facet_grid(. ~is_bot) + geom_bar() + geom_text(stat='count', aes(label=..count..), vjust=-1, size=3) + labs(title="Number of exclamation mark")
#Number of digits in screen name
ggplot(data=tweets, aes(x=num_digit_screen_name)) + facet_grid(. ~is_bot) + geom_bar() + geom_text(stat='count', aes(label=..count..), vjust=-1, size=3) + labs(title="Number of digits in screen name")
#Number of political word
ggplot(data=tweets, aes(x=num_political_word)) + facet_grid(. ~is_bot) + geom_bar() + geom_text(stat='count', aes(label=..count..), vjust=-1, size=3) + labs(title="Number of political word")
#Number of environmental word
ggplot(data=tweets, aes(x=num_environmental_word)) + facet_grid(. ~is_bot) + geom_bar() + geom_text(stat='count', aes(label=..count..), vjust=-1, size=3) + labs(title="Number of environmental word")
#Number of hashtag
ggplot(data=tweets, aes(x=num_hashtag)) + facet_grid(. ~is_bot) + geom_bar() + geom_text(stat='count', aes(label=..count..), vjust=-1, size=3) + labs(title="Number of hashtag")

bot <- tweets %>% filter(is_bot =="Bot")
no_bot <- tweets %>% filter(is_bot =="Non-Bot")

tbl_ret <- table(tweets$is_bot, tweets$include_retweet)
rownames(tbl_ret) <- c("Bot", "Non-Bot")
colnames(tbl_ret) <- c("Tweet", "Re-tweet")
print(tbl_ret)
plot(tbl_ret, cex=1.5, main="Proportion of retweet for each group")

```

## Statistical Testing   
Statistical significance between two groups, i.e. Bots and Non-Bots were tested. Average of seven numerical features were compared through two-sample t-test. One categorical feature was compared through chi-squared goodness-of-fit test.
```{r test, echo=TRUE}
#statistical testing
#Number of retweet
t1 <- t.test(log(bot$retweet+1), log(no_bot$retweet+1))

#Number of question mark
t2 <- t.test(bot$num_question, no_bot$num_question)

#Number of exclamation mark
t3 <- t.test(bot$num_exclamation, no_bot$num_exclamation)

#Number of digit in screen name
t4 <- t.test(bot$num_digit_screen_name, no_bot$num_digit_screen_name)

#Number of political word
t5 <- t.test(bot$num_political_word, no_bot$num_political_word)

#Number of environmental word
t6 <- t.test(bot$num_environmental_word, no_bot$num_environmental_word)

#Number of hashtag
t7 <- t.test(bot$num_hashtag, no_bot$num_hashtag)

#Include retweet
t8 <- chisq.test(tbl_ret)


#Create table for variables and their p-values
summary <- tweets %>% group_by(is_bot)
summary <- summary %>% summarize(Avg_Retweet = mean(retweet),
                              Avg_Question = mean(num_question),
                              Avg_Exclamation = mean(num_exclamation),
                              Avg_Digit = mean(num_digit_screen_name),
                              Avg_Political_Word = mean(num_political_word),
                              Avg_Environmental_Word = mean(num_environmental_word),
                              Avg_Hashtag = mean(num_hashtag)
                              )

#Convert gplyr summary into matrix
m <- t(as.matrix(summary[,-1])) #Remove grouping column 'is_bot'
m <- cbind(m, rep(0,7)) #Column for p-value
m <- cbind(m, rep(0,7)) #Column for Significance

colnames(m) <- c("Bot", "Non-Bot", "P-value", "Difference significant")
#Add p-values into the matrix 
m[1,3] <- t1$p.value
m[2,3] <- t2$p.value
m[3,3] <- t3$p.value
m[4,3] <- t4$p.value
m[5,3] <- t5$p.value
m[6,3] <- t6$p.value
m[7,3] <- t7$p.value

#Round value to two decimal places
m <- round(m, 2)

#Add significance into the matrix
m[1,4] <- ifelse(t1$p.value < 0.05, "Significant", "Not Significant")
m[2,4] <- ifelse(t2$p.value < 0.05, "Significant", "Not Significant")
m[3,4] <- ifelse(t3$p.value < 0.05, "Significant", "Not Significant")
m[4,4] <- ifelse(t4$p.value < 0.05, "Significant", "Not Significant")
m[5,4] <- ifelse(t5$p.value < 0.05, "Significant", "Not Significant")
m[6,4] <- ifelse(t6$p.value < 0.05, "Significant", "Not Significant")
m[7,4] <- ifelse(t7$p.value < 0.05, "Significant", "Not Significant")


#Show matrix
print(m[1:7,1:4])

#Percentage of retweets in bot tweets
retweet_pct_bot <- round(tbl_ret[1,2] / (tbl_ret[1,1] + tbl_ret[1,2]), 2)
#Percentage of retweets in non bot tweets
retweet_pct_notbot <- round(tbl_ret[2,2] / (tbl_ret[2,1] + tbl_ret[2,2]), 2)

n <- matrix(
        c(
                retweet_pct_notbot, 
                retweet_pct_bot, 
                round(t8$p.value,2), 
                ifelse(t8$p.value < 0.05, "Significant", "Not Significant")
                ),
        nrow=1,
        dimnames=list(c("Percent of retweet"), c("Not Bot", "Bot", "P-Value", "Proportion Significant")))

print(n)

# ---- R Output from hypothesis test -------------------------------------
#                       Bot      Non-Bot  P-value Difference significant
#Avg_Retweet            "422.81" "359.62" "0"     "Significant"         
#Avg_Question           "0.17"   "0.13"   "0.15"  "Not Significant"     
#Avg_Exclamation        "0.14"   "0.19"   "0.07"  "Not Significant"     
#Avg_Digit              "1.43"   "1.01"   "0"     "Significant"         
#Avg_Political_Word     "0.04"   "0.03"   "0.18"  "Not Significant"     
#Avg_Environmental_Word "0.15"   "0.05"   "0"     "Significant"         
#Avg_Hashtag            "1.35"   "2.38"   "0"     "Significant"         
#                   Not Bot Bot    P-Value Proportion Significant
#Percent of retweet "0.82"  "0.93" "0"     "Significant"  
#
# ------------------------------------------------------------------------
```

## Conclusion  
We found statistical significance in below behaviors in tweets about California wild fire.  
**Tweets by Bot are more likely to be retweeted**  
**Screen name of Bot account is more likely to have digits (Same user might have many user accounts for professional purpose)**  
**Tweets by Bot tend to comment about environmental aspects of California fire**   
**Normal tweets are more likely to embed hashtags in their tweets**  
**Tweets by Bot are more likely to be retweets of other tweets**  

This result might imply that Bot tweets are aimed at propagating specific opinion to the public by retweeting on multiple user accounts. 

## Reference
Varol, O., Ferrara, E., Davis, C., Menczer, F., & Flammini, A. (2017). Online Human-Bot Interactions: Detection, Estimation, and Characterization. Accepted paper for ICWSM'17.
